something about variable-byte encoding

### 2020-02-21 gradu/msc_english_contents.tex
minor adds

Rank and select are two succinct data structure operations which operate on a bit array B. $Rank_1$(B, i) gives the sum of 1 bits between B[0] and B[i] and 
$select_1$(b, i) gives the index of i-th 1 bit in B. Both operations work in constant time (citation?) and they require only a few percents of extra 
space over the bit array B. See Table~\ref{supportsize} for $rank$ and $select$ size requirement approximations on different sizes of bit arrays.

@@

\begin{table}
\centering
\caption{Size of support structure with SDSL-lite library}
\begin{tabular}{l||c c c c c c} 
Structure & TODO & get & multiple & size & calculations & here\\ 
\hline \hline 
$rank$ & 34.97 & 49 & 53.04 & 52.18 & 53.08 & 76.21\\
$rank_v5$ & 34.97 & 49 & 53.04 & 52.18 & 53.08 & 76.21\\
$select$ & 33.57 & 32.47 & 42.96 & 43.11 & 46.15 & 65.14\\
\hline
%
\end{tabular}
\label{supportsize}
\end{table}

@@

Simple way to calculate block length k of i-th element is to get $select_1$(i+1) and subtract previously calculated start block index from it. This however causes a second 
$select_1$ query, which is costly. A much faster option is to iterate forward in bit array until next 1 is found. An even faster option is to read an integer from 
the bit array, fix it's offset and count trailing zeros. Traveling the bit array is preferred, since it is very simple to implement and is not noticably slower. See 
Table~\ref{endblockcall} for detailed times. See Figure~\ref{select_pseudo} for a DAC with $select$ and block size 8. With different block sizes, extra calculation to get 
the block location from the byte array is needed. In the example, $CalculateLength$ returns the length of the block and $ByteMask(k)$ returns a bit mask for k bytes. $number$
is a long unsigned integer with a length of multiple bytes.

@@

The following experiments are run on <at the moment my work Lenovo Thinkpad T480s, TODO run on a server and write specs here>.

Five different kinds of VB decoding algorithms were used. Bris-implementations are from \citep{Bri09}. Rank and select functions are from
 \citep{gbmp2014sea}.

\begin{itemize}
  \item Bris4v5 - Bris implementation with block size 4, using rank support v5
  \item Bris8 - Bris implementation with block size 8, using rank support v
  \item Bris8v5 - Bris implementation with block size 8, using rank support v5
  \item Our4 - Our implementation with block size 4, using select support mcl
  \item Our8 - Our implementation with block size 8, using select support mcl
\end{itemize}

>>>>

\section{Rank and Select}
Rank and select are two succinct data structure operations which operate on a bit array B. $Rank_1$(B, i) gives the sum of 1 bits between B[0] and B[i] and 
$select_1$(b, i) gives the index of i-th 1 bit in B. Both operations work in constant time (citation?). See Table~\ref{supportsize} for $rank$ and $select$ 
size requirement approximations over the bit array using \citep{gbmp2014sea}.

@@

For every 512th bit in the array, rank stores the absolute value of 1's preceding that bit in a superblock. For every 64th bit in a superblock, a relative 
count is stored. With this setup, an rank can be instantly calculated for every 64th bit. To get rank(i), superblock[i/512] and it's relative value (i\%512)/64 
are fetched. Then 1 bits from (i-i\%64) to i are counted and added to get the correct sum. For Rankv5, superblock size is 2048 and relative count is stored every
384th bit. This offers a decent tradeoff between performance and memory usage.

- explain select here
<<explain rank and select algorithms here>>

\begin{table}
\centering
\caption{Memory requirements of SDSL rank and select data structures\label{table:supportsize}}
\begin{tabular}{l||c} 
Structure & Extra size taken\\ 
\hline \hline 
Rank   & \text{25\% of bit array} \\
Rankv5 & \text{6.25\% of bit array}\\
Select & \text{8.3\% of bit array}\\
TODO & see if data quality or size matters here\\

@@

Simple way to calculate block length k of i-th element is to get $select_1$(i+1) and subtract previously calculated start block index from it. This however causes a second 
$select_1$ query, which is costly. A much faster option is to iterate forward in bit array until next 1 is found. An even faster option is to read an integer from 
the bit array, fix it's offset and count trailing zeros. Traveling the bit array is preferred, since it is very simple to implement and is not noticably slower. See 
Table~\ref{endblockcall} for detailed times. See Figure~\ref{select_pseudo} for a DAC with $select$ and block size 8. With different block sizes, extra calculation to get 
the block location from the byte array is needed. In the example, $CalculateLength$ returns the length of the number in blocks and $ByteMask(k)$ returns a bit mask for k bytes.

@@

The following experiments are run on <at the moment my work Lenovo Thinkpad T480s, TODO run on a server and write specs here>. The algorithms 
were implemented with C++ (TODO: list compile tags and whatnot here) using data structures and functions from \citep{gbmp2014sea}.

Five different kinds of VB decoding algorithms were used. Bris-implementations are from \citep{Bri09}. 
\begin{itemize}
  \item Bris4v5 - Bris implementation with block size 4, using rank support v5
  \item Bris8 - Bris implementation with block size 8, using rank support v
  \item Bris8v5 - Bris implementation with block size 8, using rank support v5
  \item Our4 - Our implementation with block size 4, using select support mcl
  \item Our8 - Our implementation with block size 8, using select support mcl
\end{itemize}
